---
title: Malaria's effect on loan repayment in Sub-Saharan Africa
author: A PROPOSAL
csl: elsevier-harvard.csl
date: ''
output:
  pdf_document:
    toc: no
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---


```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(knitr)
library(RColorBrewer)
library(sp)
library(maptools)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

# This package
library(malaria)
```


```{r, fig.align='center', fig.height = 1.5}
# Map of Africa
par(mar = c(0, 0, 0, 0))
par(oma = c(0,0, 0, 0))
africa <- malaria::africa
africa_small <- thinnedSpatialPoly(SP = africa,
                                     minarea = 0,
                                     tolerance = 20,
                                     topologyPreserve = TRUE)
cols <- colorRampPalette(brewer.pal(9, 'Blues')[3:9])(nrow(africa_small))
cols <- sample(cols)
cols <- adjustcolor(cols, alpha.f = 0.8)
plot(africa_small, col = cols, border = NA)
```



# Context


The robust association between population and economic health is no secret. In a classic feedback loop, changes in a country's economic circumstance lead to changes in health outcomes, and vice-versa [@World1999, @Bloom2008]. However, most analyses into this association are at the population-level, since individual-level financial data has historically been scarce, small, unreliable and difficult to access. 


In the case of malaria - the most deadly disease in history - we understand the magnitude of its burden in health terms. And rough analyses suggest that malaria makes a country 33% poorer, even after adjustment for confounders [@Sachs2002]. We also know that access to microfinance is associated with improved anti-malaria behaviors [@Tarozzi2014]. But at the individual-level, is there also a feedback loop? Does a reduction in malaria lead to improved financial behavior? 


# The opportunity

Lendable has amassed granular loan repayment data on more than a quarter million individuals living in malaria-endemic zones. These data can be compared to publicly available _Plasmodium falciparum_ incidence data [@Bhatt2015] to estimate a simple model for the effect of malaria on individual-level financial behavior. 

This analysis will be relatively straightforward. But the results could be impactful since this will be the first malaria/finance study of its kind in terms of granularity and magnitude of data. 

# The case for pursuing this 

What follows is a brief list of reasons why pursuing this is a good idea.

- We have the data already. Why not? 
- Because of the uniqueness of our data, there is a high likelihood that we could publish this in a high-impact economics journal (which means good publicity, connections to academia, etc.). 
- By doing a bit of publicity around the analysis, we position ourselves as thought-leaders in this field, and draw attention from impact-oriented organizations particularly. 
- Joe has to spend a lot of his time doing this kind of stuff (publishing on malaria and economics) anyway over the next few months for his PhD (on his own dime/time). Using Lendable data is a win-win: Joe gets access to a unique dataset, Lendable gets free analysis. In other words, **there is no opportunity cost in terms of labor or capacity**.
- There could be some interesting data side-products (predictive rainfall model, elevation and land-cover analysis, etc.) that would be carried out as part of this analysis, and useful to Lendable thereafter. 

# The case against pursuing this 

- This probably won't lead to any immediate, quantifiable increases in revenue for Lendable. 
- Are there data protection / disclosure issues? 




```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown'
), 'packages.bib')
```


# References

